---
title: "nicethings"
author: "Simon Thelwall"
date: "`r nicethings::yaml_today()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{nicethings}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# nicethings

This package draws together a bunch of functions I've written to facilitate nicely formatting values calculated in R for use in Rmarkdown reports.

Using values calculated in R for printing can be long-winded. 
Often to get values into a format that is nice to read involves multiple nested functions. 
This makes inline code difficult to write and difficult to read. 
`nicethings` aims to solve this by providing simple wrappers for common tasks. 

```{r, echo=FALSE}
library(nicethings)
```

## Regression output

### The problem

Consider a regression model, from which we want to print the output into a paragraph. 

```{r, echo=TRUE}
library(broom)
data("mtcars")
m1 <- lm(mpg ~ cyl + wt, data = mtcars)
m1_tidy <- broom::tidy(m1, conf.int = TRUE)
m1_tidy
```

Say we want to print the estimate for the effect of weight on the mpg, with the 95% confidence intervals. 
To do this using base R the code might look something like this:

> Vehicle weight was significantly associated with decreased fuel efficiency `r nice_inline_r("paste0(round(m1_tidy[3,2], 1), \" (95% CI:\", round(m1_tidy[3,6], 1), \"-\", round(m1_tidy[3,7], 1), \")\")")`

which would render as: 

> Vehicle weight was significantly associated with decreased fuel efficiency `r paste0(round(m1_tidy[3,2], 1), " (95% CI:", round(m1_tidy[3,6], 1), "-", round(m1_tidy[3,7], 1), ")")`. 

This is true of various things in R: percentages, p-values, for example.

### A solution

A nicer approach is to wrap these up into a function and do something like: 

> Vehicle weight was significantly associated with decreased fuel efficiency `r nice_inline_r("nice_estimate(m1_tidy[3,2], m1_tidy[3,6], m1_tidy[3,7])")`

which would render as:

> Vehicle weight was significantly associated with decreased fuel efficiency `r nice_estimate(m1_tidy[3,2], m1_tidy[3,6], m1_tidy[3,7])`
